Structural Divergence Analysis

### 2017-04-16 analysis-structure-core.Rmd
paper y dinámica

```{r}
# read experimental data obtained from ConsurfDB
consurf = read.csv(score.fname, sep = ";")
score = consurf$SCORE[indexes.core]

>>>>

### Sequence comparison
In this section we compare theoretical sequence profiles with sequence scores obtained from ConsurfDB. To obtain theoretical profiles, we calculated the number of times we had mutated each site for each selection regimen.

```{r sequence-profiles-comparison}
# read experimental data obtained from ConsurfDB
consurf = read.csv(score.fname, sep = ";")
score = consurf$SCORE[indexes.core]

### 2017-04-16 analysis-structure-normal-modes.Rmd
paper y dinámica

```{r set_my_colors}
# set my colors
myColors = c("blue", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "strong", "mut+sel")
```

@@

## profiles to compare
### Pn
Pn.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.Pn.csv", sep = ""))
Pn.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))
Pn.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))

### energy
energy.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.va.csv", sep = ""))
energy.strong.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_strong.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))
energy.mut.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))

@@

Pn.strong = read.csv(Pn.strong.fname, header = T)
rownames(Pn.strong) = paste("strong", rownames(Pn.strong), sep = "")
energy.strong = read.csv(energy.strong.fname, header = T)
rownames(energy.strong) = paste("strong", rownames(energy.strong), sep = "")

@@

```{r prepare-mode-data}
mode = seq(ncol(Pn.strong))

# Prepare modeComparison.strong
tPn.strong = as.data.frame(t(as.matrix(Pn.strong)))
tPn.strong = cbind(data.frame("mode" = mode), tPn.strong)
Pn.strong.long = melt(tPn.strong, "mode", variable.name = "protein", value.name = "Pn")

tenergy.strong = as.data.frame(t(as.matrix(energy.strong)))
tenergy.strong = cbind(data.frame("mode" = mode), tenergy.strong)
energy.strong.long = melt(tenergy.strong, "mode", variable.name = "protein", value.name = "energy")
energy.strong.long$sn2 = 1/energy.strong.long$energy

modeComparison.strong = merge(energy.strong.long, Pn.strong.long, by = c("mode", "protein"))
modeComparison.strong = ddply(modeComparison.strong, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

@@

# modeComparison.exp has NA's because for proteins with gaps when using Keff there're less modes
#print("Warning: modeComparison.exp <- na.omit(modeComparison.exp)")
modeComparison.exp <- na.omit(modeComparison.exp)
modeComparison.strong <- na.omit(modeComparison.strong)
modeComparison.mut <- na.omit(modeComparison.mut)
```

@@

# fit median and quantiles to simulated data and add them to data.frames
fit.strong.rq.5 <- rq(Pn ~ sn2, tau = .5, data = modeComparison.strong, method = "fn")
fit.strong.rq.05 <- rq(Pn ~ sn2, tau = .05, data = modeComparison.strong, method = "fn")
fit.strong.rq.95 <- rq(Pn ~ sn2, tau = .95, data = modeComparison.strong, method = "fn")
modeComparison.strong$rq.5 = fit.strong.rq.5$fitted.values
modeComparison.strong$rq.05 = fit.strong.rq.05$fitted.values
modeComparison.strong$rq.95 = fit.strong.rq.95$fitted.values

@@

modeComparison.all.datasets = rbind(data.frame("dataset" = "exp", modeComparison.exp),
                                    data.frame("dataset" = "mut", modeComparison.mut),
                                    data.frame("dataset" = "strong", modeComparison.strong))

modeComparison.all.datasets$dataset = factor(modeComparison.all.datasets$dataset, levels = c("exp", "mut", "strong"))
```
### Pn vs. mode energy

@@

# Plot Pn vs energy
dat = modeComparison.all.datasets
dat$dataset = revalue(dat$dataset,c("strong" = "mut+sel"))

@@

p.facet = p + facet_grid(dataset~., switch = "y") + theme(legend.position = "none")
p.facet %+% subset(dat, dataset == "exp")
p.facet %+% subset(dat, dataset == "mut")
p.facet %+% subset(dat, dataset == "mut+sel")
p.facet
```

@@

dat = modeComparison.all.datasets
dat$dataset = revalue(dat$dataset,c("strong" = "mut+sel"))

@@

p.facet = p + facet_grid(dataset~., switch = "y") + theme(legend.position = "none")
p.facet %+% subset(dat,dataset == "exp")
p.facet %+% subset(dat,dataset == "mut")
p.facet %+% subset(dat,dataset == "mut+sel")
p.facet
p + theme(legend.position = "bottom")

@@

```{r correlation-mode-profiles}
dat = modeComparison.all.datasets
#dat$dataset = revalue(dat$dataset,c("strong" = "mut+sel"))

@@

dat = subset(dat, mode < 101)
dat.exp = subset(dat, dataset == "exp")
dat.mut = subset(dat,dataset == "mut")
dat.strong = subset(dat,dataset == "strong")

r.exp.mut = cor(dat.exp$Pn.median, dat.mut$Pn.median)
r.exp.strong = cor(dat.exp$Pn.median, dat.strong$Pn.median)
r.mut.strong = cor(dat.mut$Pn.median, dat.strong$Pn.median)
cor.df = data.frame("comparison" = c("exp vs mut", "exp vs strong", "mut vs strong"),
                    "R" = c(r.exp.mut, r.exp.strong, r.mut.strong))
knitr::kable(cor.df, digits = 2)

@@

plot(x = dat.exp$Pn.median, 
     y = dat.strong$Pn.median, 
     xlab = "exp.Pn.median", 
     ylab = "strong.Pn.median", 
     xlim = c(min(c(dat.exp$Pn.median, dat.strong$Pn.median)), max(c(dat.exp$Pn.median, dat.strong$Pn.median))),
     ylim = c(min(c(dat.exp$Pn.median, dat.strong$Pn.median)), max(c(dat.exp$Pn.median, dat.strong$Pn.median))))
abline(0, 1)

plot(x = dat.mut$Pn.median, 
     y = dat.strong$Pn.median, 
     xlab = "mut.Pn.median", 
     ylab = "strong.Pn.median", 
     xlim = c(min(c(dat.mut$Pn.median, dat.strong$Pn.median)), max(c(dat.mut$Pn.median, dat.strong$Pn.median))),
     ylim = c(min(c(dat.mut$Pn.median, dat.strong$Pn.median)), max(c(dat.mut$Pn.median, dat.strong$Pn.median))))
abline(0, 1)
```

>>>>

```{r set_my_colors}
# set my colors
myColors = c("blue", "darkgreen", "darkgreen", "red", "red") 
names(myColors) <- c("exp", "mut", "no-selection", "medium", "selection")
```

@@

## profiles to compare
### Pn
Pn.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.Pn.csv", sep = ""))
Pn.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))
Pn.mut.fname = file.path(data.dir,paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.Pn.csv", sep = ""))

### energy
energy.exp.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.exp.va.csv", sep = ""))
energy.medium.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_medium.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))
energy.mut.fname = file.path(data.dir, paste(family, "_R0_", R0, "_beta_no.sel_K.analysis_Keff_out_m.theo.va.csv", sep = ""))

@@

Pn.medium = read.csv(Pn.medium.fname, header = T)
rownames(Pn.medium) = paste("medium", rownames(Pn.medium), sep = "")
energy.medium = read.csv(energy.medium.fname, header = T)
rownames(energy.medium) = paste("medium", rownames(energy.medium), sep = "")

@@

```{r prepare-mode-data}
mode = seq(ncol(Pn.medium))

# Prepare modeComparison.medium
tPn.medium = as.data.frame(t(as.matrix(Pn.medium)))
tPn.medium = cbind(data.frame("mode" = mode), tPn.medium)
Pn.medium.long = melt(tPn.medium, "mode", variable.name = "protein", value.name = "Pn")

tenergy.medium = as.data.frame(t(as.matrix(energy.medium)))
tenergy.medium = cbind(data.frame("mode" = mode), tenergy.medium)
energy.medium.long = melt(tenergy.medium, "mode", variable.name = "protein", value.name = "energy")
energy.medium.long$sn2 = 1/energy.medium.long$energy

modeComparison.medium = merge(energy.medium.long, Pn.medium.long, by = c("mode", "protein"))
modeComparison.medium = ddply(modeComparison.medium, "protein", mutate, "Pn" = Pn / mean(Pn, na.rm = T))

@@

# modeComparison.exp has NA's because for proteins with gaps when using Keff there're less modes
#print("Warning: modeComparison.exp <- na.omit(modeComparison.exp)")
modeComparison.exp <- na.omit(modeComparison.exp)
modeComparison.medium <- na.omit(modeComparison.medium)
modeComparison.mut <- na.omit(modeComparison.mut)
```

@@

# fit median and quantiles to simulated data and add them to data.frames
fit.medium.rq.5 <- rq(Pn ~ sn2, tau = .5, data = modeComparison.medium, method = "fn")
fit.medium.rq.05 <- rq(Pn ~ sn2, tau = .05, data = modeComparison.medium, method = "fn")
fit.medium.rq.95 <- rq(Pn ~ sn2, tau = .95, data = modeComparison.medium, method = "fn")
modeComparison.medium$rq.5 = fit.medium.rq.5$fitted.values
modeComparison.medium$rq.05 = fit.medium.rq.05$fitted.values
modeComparison.medium$rq.95 = fit.medium.rq.95$fitted.values

@@

modeComparison.all.datasets = rbind(data.frame("dataset" = "exp", modeComparison.exp),
                                    data.frame("dataset" = "mut", modeComparison.mut),
                                    data.frame("dataset" = "medium", modeComparison.medium))

modeComparison.all.datasets$dataset = factor(modeComparison.all.datasets$dataset, levels = c("exp", "mut", "medium"))
```
### Pn vs. mode energy

@@

# Plot Pn vs energy
dat = modeComparison.all.datasets

@@

p.facet = p + facet_grid(dataset~., switch = "y") + theme(legend.position = "none")
p.facet %+% subset(dat, dataset == "exp")
p.facet %+% subset(dat, dataset == "mut")
p.facet %+% subset(dat, dataset == "medium")
p.facet
```

@@

dat = modeComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

@@

p.facet = p + facet_grid(dataset~., switch = "y") + theme(legend.position = "none")
p.facet %+% subset(dat,dataset == "exp")
p.facet %+% subset(dat,dataset == "no-selection")
p.facet %+% subset(dat,dataset == "selection")
p.facet
p + theme(legend.position = "bottom")

@@

```{r correlation-mode-profiles}
#dat$dataset = revalue(dat$dataset,c("medium" = "mut+sel"))
dat = modeComparison.all.datasets
dat$dataset = revalue(dat$dataset, c("medium" = "selection"))
dat$dataset = revalue(dat$dataset, c("mut" = "no-selection"))

@@

dat = subset(dat, mode < 101)
dat.exp = subset(dat, dataset == "exp")
dat.mut = subset(dat,dataset == "no-selection")
dat.medium = subset(dat,dataset == "selection")

r.exp.mut = cor(dat.exp$Pn.median, dat.mut$Pn.median)
r.exp.medium = cor(dat.exp$Pn.median, dat.medium$Pn.median)
r.mut.medium = cor(dat.mut$Pn.median, dat.medium$Pn.median)
cor.df = data.frame("comparison" = c("exp vs mut", "exp vs medium", "mut vs medium"),
                    "R" = c(r.exp.mut, r.exp.medium, r.mut.medium))
knitr::kable(cor.df, digits = 2)

@@

plot(x = dat.exp$Pn.median, 
     y = dat.medium$Pn.median, 
     xlab = "exp.Pn.median", 
     ylab = "medium.Pn.median", 
     xlim = c(min(c(dat.exp$Pn.median, dat.medium$Pn.median)), max(c(dat.exp$Pn.median, dat.medium$Pn.median))),
     ylim = c(min(c(dat.exp$Pn.median, dat.medium$Pn.median)), max(c(dat.exp$Pn.median, dat.medium$Pn.median))))
abline(0, 1)

plot(x = dat.mut$Pn.median, 
     y = dat.medium$Pn.median, 
     xlab = "mut.Pn.median", 
     ylab = "medium.Pn.median", 
     xlim = c(min(c(dat.mut$Pn.median, dat.medium$Pn.median)), max(c(dat.mut$Pn.median, dat.medium$Pn.median))),
     ylim = c(min(c(dat.mut$Pn.median, dat.medium$Pn.median)), max(c(dat.mut$Pn.median, dat.medium$Pn.median))))
abline(0, 1)
```

### 2017-04-16 analysis-structure.Rmd
paper y dinámica

ggplot(d, aes(x = site, y = mean.z.RSD, col = dataset)) +
  geom_line() +
  geom_text(data = subset(d, dataset == "exp" & mean.z.RSD < -.5), aes(x = site, y = mean.z.RSD, label = site), nudge_y = -.1, size = 1.8)+
  geom_vline(xintercept = active.sites, show.legend = F) +
  geom_vline(xintercept = contacts.active.sites, col = "grey", show.legend = F) + 
  scale_colour_manual(name = NULL, values = myColors)

@@

## experimental
exp.mean.z.RSD.vec <- vec2resno(as.vector(d.exp$mean.z.RSD), sort(rep(1:ncol(pdb.xyz.ca),3)))
write.pdb(xyz = c(pdb.xyz.ca), b = exp.mean.z.RSD.vec, file =  file.path(data.dir, paste(p.ref, "_exp.pdb", sep = "")))

## mut
mut.mean.z.RSD.vec <- vec2resno(as.vector(d.mut$mean.z.RSD), sort(rep(1:ncol(pdb.xyz.ca),3)))
write.pdb(xyz = c(pdb.xyz.ca), b = mut.mean.z.RSD.vec, file = file.path(data.dir, paste(p.ref, "_mut.pdb", sep = "")))

## mut + sel / medium
medium.mean.z.RSD.vec <- vec2resno(as.vector(d.medium$mean.z.RSD), sort(rep(1:ncol(pdb.xyz.ca),3)))
write.pdb(xyz = c(pdb.xyz.ca), b = medium.mean.z.RSD.vec, file = file.path(data.dir, paste(p.ref, "_medium.pdb", sep = "")))
```

### sequence comparison
```{r sequence-profiles-comparison}
# read experimental data obtained from ConsurfDB
consurf = read.csv(score.fname, sep = ";")
score = consurf$SCORE

@@

```
### sequence - structure comparison

>>>>

ggplot(d, aes(x = site, y = mean.z.RSD, col = dataset)) +
  geom_line() +
  geom_text(data = subset(d, dataset == "exp" & mean.z.RSD < -.5), aes(x = site, y = mean.z.RSD, label = site), nudge_y = -.1, size = 1.8) +
  labs(x = "site", y = "<zRSDi>") +
  geom_vline(xintercept = active.sites, show.legend = F) +
  geom_vline(xintercept = contacts.active.sites, col = "grey", show.legend = F) + 
  scale_colour_manual(name = NULL, values = myColors)

@@

## experimental
exp.mean.z.RSD.vec <- vec2resno(as.vector(d.exp$mean.z.RSD), 1:ncol(pdb.xyz.ca))
write.pdb(xyz = c(pdb.xyz.ca), b = exp.mean.z.RSD.vec, file =  file.path(data.dir, paste(p.ref, "_exp.pdb", sep = "")))

## mut
mut.mean.z.RSD.vec <- vec2resno(as.vector(d.mut$mean.z.RSD), 1:ncol(pdb.xyz.ca))
write.pdb(xyz = c(pdb.xyz.ca), b = mut.mean.z.RSD.vec, file = file.path(data.dir, paste(p.ref, "_mut.pdb", sep = "")))

## mut + sel / medium
medium.mean.z.RSD.vec <- vec2resno(as.vector(d.medium$mean.z.RSD), 1:ncol(pdb.xyz.ca))
write.pdb(xyz = c(pdb.xyz.ca), b = medium.mean.z.RSD.vec, file = file.path(data.dir, paste(p.ref, "_medium.pdb", sep = "")))
```

### Sequence comparison

In this section we compare theoretical sequence profiles with sequence scores obtained from ConsurfDB. To obtain theoretical profiles, we calculated the number of times we had mutated each site for each selection regimen.

```{r sequence-profiles-comparison}
# read experimental data obtained from ConsurfDB
consurf = read.csv(score.fname, sep = ";")
score = consurf$SCORE

@@

### Sequence - structure comparison
